A Tale of Two Transects

### 2020-02-02 writing/transect_paper.Rmd
cleo

venn_ASV %>%
  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
  set_names(gsub("_", " ", names(.))) %>%
  kable(
    booktabs = TRUE,
    caption = "(ref:venn-table-asv)",
    linesep = c(rep("", 2), "\\addlinespace", rep("", 2), "\\addlinespace"),
    align = ("rcrrcrrcrr"),
    row.names = FALSE
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
      names(venn_ASV)[-(1)] %>% 
        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
        rle() %$%
        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
      )
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
       names(venn_ASV)[-(1)] %>% 
        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
        rle() %$%
        set_names(lengths, values)
      )
  )
```

@@

```{r venn, fig.cap = "(ref:venn)", out.height = "30%", fig.show="hold"}
vennplot_ASV <- venn_ASV %>%
  rownames_to_column("set") %>%
  mutate(set = map_chr(set,
                       ~ strsplit(., "")[[1]] %>%
                     as.integer() %>%
                     as.logical() %>%
                     magrittr::extract(LETTERS[seq_along(.)], .) %>%
                     paste(collapse = "&"))) %>%
  mutate(
    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057,
                       sep = "/") %>%
      gsub("/+", "/", .) %>%
      gsub("(^/|/$)", "", .),
    label = paste(ASVs, reads_frac, sep = "\n")
  )

eulerr::venn(deframe(select(vennplot_ASV, set, ASVs))) %>%
  plot(
    legend = list(
      labels = c("PacBio Long", "PacBio Short", "Ion Torrent")
    ),
    quantities = vennplot_ASV$label,
    main = "ASV"
  )

vennplot_OTU <- venn_OTU %>%
  rownames_to_column("set") %>%
  mutate(n = str_count(set, "1")) %>%
  arrange(n, desc(set)) %>%
  mutate(set = map_chr(set,
                       ~ strsplit(., "")[[1]] %>%
                     as.integer() %>%
                     as.logical() %>%
                     magrittr::extract(LETTERS[seq_along(.)], .) %>%
                     paste(collapse = "&"))) %>%
  mutate(
    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057,
                       sep = "/") %>%
      gsub("/+", "/", .) %>%
      gsub("(^/|/$)", "", .),
    label = paste(OTUs, reads_frac, sep = "\n")
  )

@@

read_comparison <- function(multi_table, comparisons, type) {
  multi_table <- filter(multi_table, type == !!type)
  comparisons <- filter(comparisons, type == !!type)
  p <- multi_table %>%
    filter(x > 0, y > 0) %>%
    ggplot(aes(y = y, x = x)) +
    geom_point(alpha = 0.2, shape = 1) +
    scale_x_log10() +
    scale_y_log10(limits = c(1, 1e4)) +
    coord_equal() +
    xlab(NULL) +
    ylab(NULL) +
    geom_rug(aes(y = y), sides = "l", alpha = 0.2,
             data = multi_table %>% filter(x == 0, y > 0)) +
    geom_rug(aes(x = x), sides = "b", alpha = 0.2,
             data = multi_table %>% filter(y == 0, x > 0)) +
    facet_grid(y_var ~ x_var, switch = "both", space = "free", ) +
    stat_smooth(method = "loess", formula = y ~ x) +
    geom_abline(aes(intercept = `(Intercept)`, slope = 1),
                linetype = 2L, alpha = 0.5, data = comparisons) +
    geom_text(
      aes(label = label),
      data = comparisons,
      x = 0.5,
      y = 3.5,
      color = "black",
      hjust = 0,
      parse = TRUE,
      inherit.aes = FALSE
    ) +
    theme(strip.placement = "outside", strip.background = element_blank())
  g <- ggplotGrob(p)
  g$grobs[g$layout$name %in% c("panel-1-2")] <- NULL
  g$layout <- g$layout[!(g$layout$name %in% c("panel-1-2")),]
  g
}
ggpubr::ggarrange(
  read_comparison(multi_table, comparisons, type = "ASV"),
  read_comparison(multi_table, comparisons, type = "OTU"),
  ncol = 1,
  labels = "auto",
  label.x = 0.9
)

@@

```{r taxa, fig.cap = "(ref:taxa)", out.width="100%", fig.width = 8}
set.seed(4)
readd(taxdata, cache = cache) %>%
  metacoder::heat_tree(
    layout = "davidson-harel",
    # initial_layout = "reingold-tilford",
    node_size = rowMeans(.$data$tax_read[,-1]),
    node_color = rowMeans(.$data$tax_read[,-1]),
    node_label = replace_na(taxon_names, "unknown"),
    node_size_range = c(0.005, .035),
    node_size_axis_label = "Read abundance",
    node_color_axis_label = "Read abundance",
    node_label_size_range = c(0.01, 0.03),
    edge_size = rowMeans(.$data$tax_asv[,-1]),
    edge_size_trans = "linear",
    edge_size_axis_label = "ASV count",
    edge_size_range = c(0.0025, 0.03),
    edge_color = rowMeans(.$data$tax_asv[,-1]),
    edge_color_axis_label = "ASV count",
    # aspect_ratio = 3/2,
    make_node_legend = FALSE,
    make_edge_legend = FALSE,
    output_file = here("temp/heattree.pdf")
  )
```

@@

light_plan %>%
  filter(startsWith(target, "correlog")) %>%
  select(target) %>%
  separate(
    target,
    into = c("step", "guild", "metric", "tech", "amplicon", "tdist"),
    sep = "_",
    remove = FALSE
  ) %>%
  mutate(
    tdist = paste(tdist, "year"),
    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
    amplicon = factor(amplicon, levels = c("short", "long"), labels = c("short amplicon", "long amplicon"))
  ) %>%
  mutate_at("target", purrr::map, readd, character_only = TRUE, cache = cache) %>%
  mutate_at("target", purrr::map, "mantel.res") %>%
  mutate_at("target", purrr::map, as.data.frame) %>%
  unnest(target) %>%
  filter(complete.cases(.)) %>%
  mutate(Significant = `Pr(corrected)` < 0.05) %>%
  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
  geom_point(size = 3) +
  geom_line(alpha = 0.3, size = 1.5) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  # geom_line(aes(y = exp(predict(bc_fit)))) +
  scale_y_continuous(limits = c(-0.05, NA), name = "Mantel correlation") +
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12)) +
  scale_color_discrete(name = "Time lag") +
  xlab("Distance (m)") +
  facet_grid(metric + amplicon ~ guild, scales = "fixed") +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1))
```

@@

variog_empir %>%
  ggplot(aes(dist, gamma, color = timelag, size = n)) +
  # geom_hline(data = variog_sills, aes(yintercept = sills), linetype = "dashed", color = "gray50") +
  # geom_vline(data = variog_data, aes(xintercept = range), linetype = "dashed", color = "gray50") +
  # geom_jitter(shape = 46, alpha = 0.5) +
  # geom_line(size = 1.5, alpha = 0.5) +
  geom_point(alpha = 0.5) +
  # geom_smooth(data = variog_points, aes(fill = timelag), color = FALSE, alpha = 0.5) +
  geom_line(data = filter(variog_fit, timelag == 1), size = 0.5, color = "darkcyan", alpha = 0.8) +
  geom_line(data = filter(variog_fit, timelag == 0), size = 0.5, color = "red1", alpha = 0.8) +
  scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
  scale_color_discrete(name = "Time lag") +
  scale_y_continuous(name = "Community dissimilarity") +
  facet_grid(metric + amplicon ~ guild, scales = "free")

@@

```{r variofit}
variog_data %>%
  mutate_at("variofitST", lapply, purrr::compose(as.data.frame, as.list, gstat::extractPar)) %>%
  unnest(variofitST) %>%
  arrange(metric, amplicon, guild) %>%
  select(Guild = guild, Amplicon = amplicon, Metric = metric,
         Range = range, Anisotropy = anis) %>%
  mutate(Range = 3*Range,
         "Time Range" = Range/Anisotropy,
         "Time Range" = ifelse(`Time Range` > 100, NA, `Time Range`)) %>%
  mutate_at(c("Range", "Time Range"), formatC, digits = 1,
            format = "f") %>%
  mutate_at("Range", paste, "m") %>%
  mutate_at("Time Range", paste, "a") %>%
  mutate_at("Time Range", str_replace, "NA a", "-") %>%
  mutate_at("Amplicon", fct_relabel, str_replace, " amplicon", "") %>%
  select(Guild, Amplicon, Metric, "Space Range" = Range, "Time Range") %>%
  kable(booktabs = TRUE, caption = "(ref:variofit)")
```

@@

```{r full-length, fig.cap = "(ref:full-length)", fig.height = 6}
full_dens <- reads_table %>%
  dplyr::left_join(dplyr::select(datasets, seq_run, tech)) %>%
  dplyr::filter(region %in% c("long", "short")) %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate(
    amplicon = ifelse(seq_run == "pb_500", "Long", "Short") %>%
      factor(levels = c("Short", "Long")),
    reads = reads / sum(reads)
  ) %>%
  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
  stat_density(bw = 0.5, geom = "line", position = "identity") +
  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
  scale_y_continuous(name = "Density") +
  scale_x_continuous(name = NULL) +
  facet_wrap(~tech + amplicon, ncol = 1, strip.position = "right", scales = "free_y")

full_ecdf <- reads_table %>%
  dplyr::filter(region  %in% c("long", "short")) %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate(reads = reads / sum(reads)) %>%
  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
  stat_ecdf(geom = "line") +
  scale_color_brewer(type = "qual", labels = c("Short amplicon - Ion", "Short amplicon - PacBio", "Long amplicon - PacBio"), name = NULL, palette = "Set2") +
  scale_y_continuous(name = "Cumulative read fraction") +
  scale_x_continuous(name = "Length (bp)") +
  theme(legend.position = c(0.98, 0.02), legend.justification = c(1, 0))

@@

```{r regions, fig.cap="(ref:region-asv)"}
region_table %>%
  dplyr::filter(seq_run == "pb_500") %>%
  ggplot(aes(reads, ASVs, color = length_med, label = region)) +
  geom_point() +
  ggrepel::geom_text_repel() +
  scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
  xlab("Mapped reads")
```

@@

reads_table %>%
  dplyr::filter(region %in% baseregions,
         seq_run == "pb_500") %>%
  dplyr::mutate_at(
    "region",
    factor,
    levels = baseregions
    ) %>%
  ggplot(aes(length, weight = reads)) +
  stat_density(bw = 0.5, color = "#8da0cb", geom = "line") +
  geom_vline(data = region_limits,
             aes(xintercept = length),
             linetype = "dashed", alpha = 0.5) +
  facet_wrap(~region,ncol = 1, strip.position = "right", scales = "free_y") +
  scale_y_continuous(breaks = NULL, labels = NULL, name = "Read density") +
  scale_x_continuous(limits = c(0, 500), name = "Length (bp)")
```

@@

```{r ITS2-length, fig.cap = "(ref:ITS2-length)", fig.height = 6}
its2_dens <- reads_table %>%
  dplyr::left_join(dplyr::select(datasets, seq_run, tech)) %>%
  dplyr::filter(region == "ITS2") %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate(
    amplicon = ifelse(seq_run == "pb_500", "Long", "Short") %>%
      factor(levels = c("Short", "Long")),
    reads = reads / sum(reads)
  ) %>%
  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
  stat_density(bw = 0.5, geom = "line", position = "identity") +
  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
  scale_y_continuous(name = "Density") +
  scale_x_continuous(name = NULL) +
  facet_wrap(~tech + amplicon, ncol = 1, strip.position = "right", scales = "free_y")

its2_ecdf <- reads_table %>%
  dplyr::filter(region == "ITS2") %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate(reads = reads / sum(reads)) %>%
  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
  stat_ecdf(geom = "line") +
  scale_color_brewer(type = "qual", labels = c("Short amplicon - Ion", "Short amplicon - PacBio", "Long amplicon - PacBio"), name = NULL, palette = "Set2") +
  scale_y_continuous(name = "Cumulative read fraction") +
  scale_x_continuous(name = "ITS2 Length (bp)") +
  theme(legend.position = c(0.98, 0.02), legend.justification = c(1, 0))

@@

set.seed(4)
taxdata2 %>%
  metacoder::heat_tree(
    layout = "davidson-harel",
    # initial_layout = "reingold-tilford",

@@

@@

    node_size = rowMeans(.$data$tax_read[,-1]),
    node_size_range = c(0.01, .035),
    node_size_axis_label = "Read abundance",

@@

    edge_size = rowMeans(.$data$tax_asv[,-1]),
    edge_size_range = c(0.005, 0.03),
    edge_size_trans = "linear",
    edge_size_axis_label = "ASV count",

@@

```{r sample-depth, fig.cap = "(ref:sample-depth)"}
readd(proto_physeq, cache = cache) %>% {
  mutate(
    sample_data(.),
    reads = otu_table(.) %>% rowSums()
  )
} %>%
  filter(tech != "Ion Torrent", !is.na(year), !is.na(seq_run), !is.na(x)) %>%
  complete(seq_run, tech, amplicon, site, x, nesting(year, qual)) %>%
  mutate(
    qual = fct_collapse(qual, " " = c("*", "")),
    "Sample group" = paste(year, qual)
  ) %>%
  ggplot(aes(x = x, y = reads, group = `Sample group`, fill = `Sample group`)) +
  geom_col(position = "dodge") +
  ggnomics::facet_nested(
    cols = vars(tech, amplicon, site),
    rows = vars(`Sample group`),
    nest_line = TRUE,
    resect = unit(1, "mm")
  ) +
  scale_y_log10() +
  xlab("x (m)") +
  theme(
    legend.position = "bottom",
    strip.background = element_blank()
  ) +
  geom_hline(yintercept = 100, linetype = 2, alpha = 0.5)

>>>>

venn_ASV %>%
  filter(!is.na(ASVs)) %>%
  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
  set_names(gsub("_", " ", names(.))) %>%
  kable(
    booktabs = TRUE,
    caption = "(ref:venn-table-asv)",
    linesep = c(rep("", 2), "\\addlinespace", rep("", 2), "\\addlinespace"),
    align = ("rcrrcrrcrr"),
    row.names = FALSE
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
      names(venn_ASV)[-(1)] %>% 
        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
        rle() %$%
        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
      )
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
       names(venn_ASV)[-(1)] %>% 
        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
        rle() %$%
        set_names(lengths, values)
      )
  )
```

@@

```{r venn, fig.cap = "(ref:venn)", out.height = "30%", fig.show="hold"}
vennplot_ASV <- venn_ASV %>%
  rownames_to_column("set") %>%
  mutate(set = map_chr(set,
                       ~ strsplit(., "")[[1]] %>%
                     as.integer() %>%
                     as.logical() %>%
                     magrittr::extract(LETTERS[seq_along(.)], .) %>%
                     paste(collapse = "&"))) %>%
  mutate(
    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057, `frac_SH-2257`,
                       sep = "/") %>%
      gsub("/+", "/", .) %>%
      gsub("(^/|/$)", "", .),
    label = paste(ASVs, reads_frac, sep = "\n")
  )

eulerr::venn(deframe(select(vennplot_ASV, set, ASVs)) %>% replace_na(0)) %>%
  plot(
    legend = list(
      labels = c("PacBio Long", "PacBio Short", "Ion Torrent", "Illumina")
    ),
    quantities = vennplot_ASV$label,
    main = "ASV"
  )

vennplot_OTU <- venn_OTU %>%
  rownames_to_column("set") %>%
  mutate(n = str_count(set, "1")) %>%
  arrange(n, desc(set)) %>%
  mutate(set = map_chr(set,
                       ~ strsplit(., "")[[1]] %>%
                     as.integer() %>%
                     as.logical() %>%
                     magrittr::extract(LETTERS[seq_along(.)], .) %>%
                     paste(collapse = "&"))) %>%
  mutate(
    reads_frac = paste(frac_pb_500, frac_pb_483, frac_is_057, `frac_SH-2257`,
                       sep = "/") %>%
      gsub("/+", "/", .) %>%
      gsub("(^/|/$)", "", .),
    label = paste(OTUs, reads_frac, sep = "\n")
  )

@@

read_comparison <- function(multi_table, comparisons, type) {
  multi_table <- filter(multi_table, type == !!type)
  comparisons <- filter(comparisons, type == !!type)
  p <- multi_table %>%
    filter(x > 0, y > 0) %>%
    ggplot(aes(y = y, x = x)) +
    geom_point(alpha = 0.2, shape = 1) +
    scale_x_log10() +
    scale_y_log10() +
    coord_equal() +
    xlab(NULL) +
    ylab(NULL) +
    geom_rug(aes(y = y), sides = "l", alpha = 0.2,
             data = multi_table %>% filter(x == 0, y > 0)) +
    geom_rug(aes(x = x), sides = "b", alpha = 0.2,
             data = multi_table %>% filter(y == 0, x > 0)) +
    facet_grid(y_var ~ x_var, switch = "both", space = "free") +
    stat_smooth(method = "loess", formula = y ~ x) +
    geom_abline(aes(intercept = `(Intercept)`, slope = 1),
                linetype = 2L, alpha = 0.5, data = comparisons) +
    geom_text(
      aes(label = label),
      data = comparisons,
      x = 0.2,
      y = 5.5,
      color = "black",
      hjust = 0,
      parse = TRUE,
      inherit.aes = FALSE
    ) +
    theme(strip.placement = "outside", strip.background = element_blank())
  g <- ggplotGrob(p)
  g$grobs[g$layout$name %in% c("panel-1-2", "panel-1-3", "panel-2-3")] <- NULL
  g$layout <- g$layout[!(g$layout$name %in% c("panel-1-2", "panel-1-3", "panel-2-3")),]
  g
}
ggpubr::ggarrange(
  read_comparison(multi_table, comparisons, type = "ASV"),
  read_comparison(multi_table, comparisons, type = "OTU"),
  ncol = 1,
  labels = "auto",
  label.x = 0.9
)

@@

```{r taxa, fig.cap = "(ref:taxa)", out.width="100%", fig.width = 8}
set.seed(2)
readd(taxdata, cache = cache) %>%
  metacoder::heat_tree(
    layout = "davidson-harel",
    # initial_layout = "reingold-tilford",
    node_size = rowMeans(.$data$tax_read[,-1]),
    node_color = rowMeans(.$data$tax_read[,-1]),
    node_label = replace_na(taxon_names, "unknown"),
    node_size_range = c(0.002, .035),
    node_size_axis_label = "Read abundance",
    node_color_axis_label = "Read abundance",
    node_label_size_range = c(0.01, 0.03),
    edge_size = rowMeans(.$data$tax_asv[,-1]),
    edge_size_trans = "linear",
    edge_size_axis_label = "ASV count",
    edge_size_range = c(0.001, 0.03),
    edge_color = rowMeans(.$data$tax_asv[,-1]),
    edge_color_axis_label = "ASV count",
    # aspect_ratio = 3/2,
    make_node_legend = FALSE,
    make_edge_legend = FALSE,
    output_file = here("temp/heattree.pdf")
  )
```

@@

light_plan %>%
  filter(startsWith(target, "correlog")) %>%
  select(target) %>%
  separate(
    target,
    into = c("step", "guild", "metric", "tech", "amplicon", "tdist"),
    sep = "_",
    remove = FALSE
  ) %>%
  mutate(
    tdist = paste(tdist, "year"),
    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
    amplicon = factor(amplicon, levels = c("short", "long"), labels = c("short amplicon", "long amplicon"))
  ) %>%
  mutate_at("target", purrr::map, readd, character_only = TRUE, cache = cache) %>%
  mutate_at("target", purrr::map, "mantel.res") %>%
  mutate_at("target", purrr::map, as.data.frame) %>%
  unnest(target) %>%
  filter(complete.cases(.)) %>%
  mutate(Significant = `Pr(corrected)` < 0.05) %>%
  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
  geom_point(size = 3) +
  geom_line(alpha = 0.3, size = 1.5) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  # geom_line(aes(y = exp(predict(bc_fit)))) +
  scale_y_continuous(limits = c(-0.05, NA), name = "Mantel correlation") +
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12)) +
  scale_color_discrete(name = "Time lag") +
  xlab("Distance (m)") +
  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "fixed") +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1))
```

@@

variog_empir %>%
  ggplot(aes(dist, gamma, color = timelag, size = n)) +
  # geom_hline(data = variog_sills, aes(yintercept = sills), linetype = "dashed", color = "gray50") +
  # geom_vline(data = variog_data, aes(xintercept = range), linetype = "dashed", color = "gray50") +
  # geom_jitter(shape = 46, alpha = 0.5) +
  # geom_line(size = 1.5, alpha = 0.5) +
  geom_point(alpha = 0.5) +
  # geom_smooth(data = variog_points, aes(fill = timelag), color = FALSE, alpha = 0.5) +
  geom_line(data = filter(variog_fit, timelag == 1), size = 0.5, color = "darkcyan", alpha = 0.8) +
  geom_line(data = filter(variog_fit, timelag == 0), size = 0.5, color = "red1", alpha = 0.8) +
  scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
  scale_color_discrete(name = "Time lag") +
  scale_y_continuous(name = "Community dissimilarity") +
  ggnomics::facet_nested(metric + tech + amplicon ~ guild, scales = "free")

@@

```{r variofit}
variog_data %>%
  mutate_at("variofitST", lapply, purrr::compose(as.data.frame, as.list, gstat::extractPar)) %>%
  unnest(variofitST) %>%
  arrange(metric, amplicon, tech, guild) %>%
  select(Guild = guild, Amplicon = amplicon, Tech = tech, Metric = metric,
         Range = range, Anisotropy = anis) %>%
  mutate(Range = 3*Range,
         "Time Range" = Range/Anisotropy,
         "Time Range" = ifelse(`Time Range` > 100, NA, `Time Range`)) %>%
  mutate_at(c("Range", "Time Range"), formatC, digits = 1,
            format = "f") %>%
  mutate_at("Range", paste, "m") %>%
  mutate_at("Time Range", paste, "a") %>%
  mutate_at("Time Range", str_replace, "NA a", "-") %>%
  mutate_at("Amplicon", fct_relabel, str_replace, " amplicon", "") %>%
  select(Guild, Amplicon, Tech, Metric, "Space Range" = Range, "Time Range") %>%
  kable(booktabs = TRUE, caption = "(ref:variofit)")
```

@@

```{r full-length, fig.cap = "(ref:full-length)", fig.height = 6}
full_dens <- 
  readd(reads_table, cache = cache) %>%
  dplyr::left_join(dplyr::select(datasets, seq_run, tech)) %>%
  dplyr::filter(region %in% c("long", "short")) %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate(
    amplicon = ifelse(seq_run == "pb_500", "Long", "Short") %>%
      factor(levels = c("Short", "Long")),
    reads = reads / sum(reads)
  ) %>%
  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
  stat_density(bw = 0.5, geom = "line", position = "identity") +
  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
  scale_y_continuous(name = "Density") +
  scale_x_continuous(name = NULL) +
  facet_wrap(~tech + amplicon, ncol = 1, strip.position = "right", scales = "free_y")

full_ecdf <- readd(reads_table, cache = cache) %>%
  dplyr::filter(region  %in% c("long", "short")) %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate(reads = reads / sum(reads)) %>%
  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
  stat_ecdf(geom = "line") +
  scale_color_brewer(type = "qual",
                     labels = c("Short - Ion", "Short - PacBio", "Long - PacBio", "Short - Illumina"),
                     name = NULL, palette = "Set2") +
  scale_y_continuous(name = "Cumulative read fraction") +
  scale_x_continuous(name = "Length (bp)") +
  theme(legend.position = c(1, 0.02), legend.justification = c(1, 0), legend.background = element_blank())

@@

```{r regions, fig.cap="(ref:region-asv)"}
readd(region_table, cache = cache) %>%
  dplyr::filter(seq_run == "pb_500") %>%
  ggplot(aes(reads, ASVs, color = length_med, label = region)) +
  geom_point() +
  ggrepel::geom_text_repel() +
  scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
  xlab("Mapped reads")
```

@@

readd(reads_table, cache = cache) %>%
  dplyr::filter(region %in% baseregions,
         seq_run == "pb_500") %>%
  dplyr::mutate_at(
    "region",
    factor,
    levels = baseregions
    ) %>%
  ggplot(aes(length, weight = reads)) +
  stat_density(bw = 0.5, color = "#8da0cb", geom = "line") +
  geom_vline(data = region_limits,
             aes(xintercept = length),
             linetype = "dashed", alpha = 0.5) +
  facet_wrap(~region,ncol = 1, strip.position = "right", scales = "free_y") +
  scale_y_continuous(breaks = NULL, labels = NULL, name = "Read density") +
  scale_x_continuous(limits = c(0, 500), name = "Length (bp)")
```

@@

```{r ITS2-length, fig.cap = "(ref:ITS2-length)", fig.height = 6}
its2_dens <- readd(reads_table, cache = cache) %>%
  dplyr::left_join(dplyr::select(datasets, seq_run, tech)) %>%
  dplyr::filter(region == "ITS2") %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate(
    amplicon = ifelse(seq_run == "pb_500", "Long", "Short") %>%
      factor(levels = c("Short", "Long")),
    reads = reads / sum(reads)
  ) %>%
  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
  stat_density(bw = 0.5, geom = "line", position = "identity") +
  scale_color_brewer(type = "qual", guide = "none", palette = "Set2") +
  scale_y_continuous(name = "Density") +
  scale_x_continuous(name = NULL) +
  facet_wrap(~tech + amplicon, ncol = 1, strip.position = "right", scales = "free_y")

its2_ecdf <- readd(reads_table, cache = cache) %>%
  dplyr::filter(region == "ITS2") %>%
  dplyr::group_by(seq_run) %>%
  dplyr::mutate(reads = reads / sum(reads)) %>%
  ggplot(aes(length, weight = reads, color = seq_run, group = seq_run)) +
  stat_ecdf(geom = "line") +
  scale_color_brewer(type = "qual", labels = c("Short amplicon - Ion", "Short amplicon - PacBio", "Long amplicon - PacBio"), name = NULL, palette = "Set2") +
  scale_y_continuous(name = "Cumulative read fraction") +
  scale_x_continuous(name = "ITS2 Length (bp)") +
  theme(legend.position = c(0.98, 0.02), legend.justification = c(1, 0))

@@

set.seed(2)
taxdata2 %>%
  metacoder::heat_tree(
    layout = "davidson-harel",
    # initial_layout = "reingold-tilford",

    node_size = rowMeans(.$data$tax_read[,-1]),
    node_size_range = c(0.002, .035),
    node_size_axis_label = "Read abundance",

@@

    edge_size = rowMeans(.$data$tax_asv[,-1]),
    edge_size_range = c(0.001, 0.03),
    edge_size_trans = "linear",
    edge_size_axis_label = "ASV count",

@@

    node_size = rowMeans(.$data$tax_read[,-1]),
    node_size_range = c(0.002, .035),
    node_size_axis_label = "Read abundance",

@@

    edge_size = rowMeans(.$data$tax_asv[,-1]),
    edge_size_range = c(0.001, 0.03),
    edge_size_trans = "linear",
    edge_size_axis_label = "ASV count",

@@

```{r sample-depth, fig.cap = "(ref:sample-depth)"}
readd(proto_physeq, cache = cache) %>% {
  mutate(
    sample_data(.),
    reads = otu_table(.) %>% rowSums()
  )
} %>%
  filter(tech != "Ion Torrent", !is.na(year), !is.na(seq_run), !is.na(x)) %>%
  complete(nesting(seq_run, tech, amplicon), site, x, nesting(year, qual)) %>%
  mutate(
    qual = fct_collapse(qual, " " = c("*", "")),
    "Sample group" = paste(year, qual)
  ) %>%
  ggplot(aes(x = x, y = reads, group = `Sample group`, fill = `Sample group`)) +
  geom_col(position = "dodge") +
  ggnomics::facet_nested(
    cols = vars(tech, amplicon, site),
    rows = vars(`Sample group`),
    nest_line = TRUE,
    resect = unit(1, "mm")
  ) +
  scale_y_log10() +
  xlab("x (m)") +
  theme(
    legend.position = "bottom",
    strip.background = element_blank()
  ) +
  geom_hline(yintercept = 100, linetype = 2, alpha = 0.5)
